|

### 2019-03-28 rap/article.Rmd
more text

On 11 September 1973, Army General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende. The coup marked the beginning of a seventeen-year military dictatorship which combined large-scale political repression with rapid economic liberalisation [@meadowcroft2014hayek; @valdes1995pinochet]. In 1991, after the transition to democracy, then-President Patricio Aylwin created a commission to investigate human rights abuses in the Pinochet regime. The Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], records over 2,000 cases of murders and disappearances from 1973 to 1990. In most cases, there are some details as to the personal characteristics of the victim, and where and when they died or disappeared. Despite its omissions[^omissions].

Although the Report is a valuable resource for scholars across the humanities, quantitative researchers cannot easily use the rich information it contains. In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we constructed to complement the original data. We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each of the victims, then matched the location of the human rights abuses to geographic coordinates. We matched all cases which had spatial information available.

Apart from the geographical location of the incidents, our dataset also includes: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that particular event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence. If the Report does not have a particular information, we coded it as missing. As each individual receives their own id number, new information can be added to the dataset as archival work on the Pinochet era progresses. In the following sections, we describe the dataset, show summary statistics for selected variables, then provide suggestions about how our data can answer future research questions.

@@

Our dataset comprises 2,398 observations and 57 variables. As we noted above, each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number. There are several variables describing personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available. The dataset also includes information about torture or assassination methods and geographical coordinates for a number of the incidents. Figure 01 shows an example entry from our dataset.

@@

Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values (`.csv`) file. We have also created an `R` package called `pinochet` which includes the dataset in both formats plus the codebook. The files follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy]. The `pinochet` package is available for download on the Comprehensive R Archive Network (CRAN) and at <http://github.com/danilofreire/pinochet>. To make the data accessible, our GitHub repository has detailed installation instructions for users new to `R`.

@@

The Report distinguishes between different types of violence carried out by the Pinochet regime. The first type is _deaths_. These are cases where the Commission signals a definite and known death of the victim. Take the case of Benito Heriberto Torres Torres, one of the first victims of the Pinochet regime (id number 2). Our dataset shows that Mr Torres was male, 57 years-old, and that he worked as a plumber. On the 12 September 1973, just one day after the military coup, Mr Torres was torture on the way to the 26th police station in Santiago. The records indicate he executed and his body was later found in Las Barrancas, Santiago. The dataset also shows that we obtained this information from pages 159-160 of the Truth Commission Report.

The second type of violence recorded in the dataset is _disappearances_. These are cases where government agents are presumed to have killed and disposed of the body of the victims. One such example is that of the Brazilian engineer Tulio Roberto Quintiliano Cardozo (id number 5). Mr Cardozo was a member of the Communist party and troops took him to the Military Academy for interrogation also on the 12 September 1971. He was never seen again and is presumed dead.

The third category is _disappearance, information of death_. As the name implies, these observations refer to cases the Commission confirms the individual died after being missing. The formal definition for this category is that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [@report1991 44]. The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854). He was taken in custody to the Colina air base, then loaded onto a helicopter with other political prisoners. According to the Report, government agents drugged him, beat him with a metal bar, opened his stomach with a knife before throwing the former Communist alderman from a helicopter. While his body has not been found, there is enough information about the case to classify his death as a severe human rights violation.

@@

We can also explore how the government employed violence across space. We see that all types of violence were concentrated around Santiago and neighbouring cities. The results are in line with theoretical expectations. Santiago has been Chile's most densely populated city since the colonial times and in 1975 the city was home to about 30% of the country's population [@un1976yearbook].

[comment]: <> (Figure 03 - Spatial variation. Add a few more lines above)

@@

[^df]: Postdoctoral Research Fellow, The Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io). Corresponding author.

>>>>

On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende. The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@meadowcroft2014hayek; @valdes1995pinochet]. In 1991, after Chile's transition to democracy, then-President Patricio Aylwin convened a commission to investigate human rights abuses in the Pinochet regime. The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990. The Commission directly confronted the abuses of the past and portrayed a careful reconstruction of state repression during the military government. Most importantly, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].

Although the Report is a valuable resource to scholars, quantitative researchers cannot readily use the rich information it contains. In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we collected to complement the original data. We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each victim, then georeferenced the location of the human rights abuses. We matched all cases that had spatial information available.

Apart from the geographical location of the incidents, our dataset includes variables on: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that particular event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence. If the Report does not have a particular information, we coded it as missing. As each individual receives their own id number, new information can be added to the dataset as archival work progresses. In the following sections, we describe our dataset, show summary statistics for selected variables, and suggest how our data can help answer future research questions.

@@

Our dataset comprises 2,398 observations and 57 variables. Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number. There are several variables describing the personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available. The data also feature geographical coordinates for a number of the incidents, as well as information about methods of torture and causes of death. Figure 01 shows an example entry from our dataset.

@@

Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values text file (`.csv`).  We have also created an `R` package called `pinochet` which includes the dataset in both formats plus the data codebook. The files follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy]. The `pinochet` package is available for download on the Comprehensive R Archive Network (CRAN) and at <http://github.com/danilofreire/pinochet>. To make the data accessible, our GitHub repository has detailed installation instructions for users new to `R`.

@@

The Report distinguishes between different types of violence carried out by the Pinochet regime. The first type is _deaths_. These are cases where the Commission signals a definite and known death of the victim. Take the case of Benito Heriberto Torres Torres, one of the first victims of the Pinochet regime (id number 2). Our dataset shows that Mr Torres was male, 57 years-old, and that he worked as a plumber. On the 12 September 1973, just one day after the military coup, Mr Torres was torture on the way to the 26th police station in Santiago. The records indicate he was executed and that his body was later found in Las Barrancas, Santiago. The dataset also shows that we obtained this information from pages 159-160 of the Truth Commission Report.

The second type of violence recorded in the dataset is _disappearances_. These are cases where government agents are presumed to have killed and disposed of the body of the victims. One such example is that of the Brazilian engineer Tulio Roberto Quintiliano Cardozo (id number 5). Mr Cardozo was a member of the Communist party and troops took him to the Military Academy for interrogation also on the 12 September 1973. He was never seen again and is presumed dead.

The third category is _disappearance, information of death_. As the name implies, these observations refer to cases the Commission confirms the individual died after being missing. The formal definition for this category is the following: "the victims are dead; they died at the hands of the government agents, or persons in their service; and these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [@report1991 44]. The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854). He was taken in custody to the Colina air base, then loaded onto a helicopter with other political prisoners. According to the Report, government agents drugged him, beat him with a metal bar, opened his stomach with a knife before throwing the former Communist alderman from a helicopter. While his body has not been found, there is enough information about the case to classify his death as a severe human rights violation.

@@

We can also explore how the government employed violence across space. We see that assassinations are concentrated around Santiago and neighbouring cities. The results are expected. Santiago has been Chile's most densely populated city since the colonial times and in 1975 the city was home to about 30% of the country's population [@un1976yearbook]. Disappearances follow a similar pattern. Although there are cases in the northern parts of the country, the military government targeted more victims in Santiago and neighbouring areas.

[comment]: <> (Figure 03 - Spatial variation)

We can further disaggregate the analysis and visualise how the regime persecuted particular groups. Students and members of left-wing parties are two categories of special interest. On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium. Military forces tortured and murdered hundreds of prisoners over the course of a few days. An emblematic victim was Víctor Lidio Jara Martínez, a university teacher and song composer. According to witnesses, Mr Jara had his hand crushed and his wrists broken, and he was later shot 44 times with a machine gun. Mr Jara's body was left in the front of the stadium as a warning for other political prisoners [@mcsherry2015victor 56]. The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR). The MIR originated in student unions and in the late 1970s the movement promoted a series of attacks against government personnel and official buildings. Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].

[comment]: <> (Figure 04 - Students and Communist Leaders)

@@

[^df]: Postdoctoral Research Fellow, the Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io). Corresponding author.
